---
title: "Iteration and listcols"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(rvest)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuious.color = "viridis",
  ggplot2.continuious.fill = "viridis"
  
)

scale_color_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


Set seed for reproducibility: 
```{r}
set.seed(12345)
```


# Lists

```{r}
vec_numeric = 1:4
vec_char = c("my", "name", "is", "vicky")

tibble(
  num = vec_numeric,
  char = vec_char
)
```

Different stuff with different lengths

```{r}
l = list(
    vec_numeric = 1:5,
    vec_char = LETTERS,
    matrix(1:10, nrow = 5, ncol = 2),
    summary = summary(rnorm(100))
)

```

Accessing lists: 

```{r}
l$vec_char # <-- extracting the character vector
l[[1]] # <-- shows you what the 1st element in the list is (replace w diff numbers to see different elements)
l[["summary"]] 
```


# Loops!

```{r}
list_norm_samples = 
  list(
    a = rnorm(20, 1, 5),
    b = rnorm(20, 0, 7),
    c = rnorm(20, 20, 5),
    d = rnorm(20, -45, 13)
  )
```

Using the mean and sd function from functions unit: 

```{r}
mean_and_sd = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric")
  } else if (length(x) == 1) {
    stop("Cannot be computed for length 1 vectors")
  }
  
  mean_x = mean(x)
  sd_x = sd(x)

   tibble(
    mean = mean_x, 
    sd = sd_x
  )
}
```

Applying the function 
```{r}
mean_and_sd(list_norm_samples$a)
mean_and_sd(list_norm_samples$b)
mean_and_sd(list_norm_samples$c)
mean_and_sd(list_norm_samples$d)

```

It is easier to apply the function when you use a loop function or a for loop to keep track of all the elements of your input list and saves the results of applying the function to the output list 

```{r}
output = vector("list", length = 4)

for (i in 1:4) {
  output[[i]] = mean_and_sd(list_norm_samples[[i]])
  
}
output # <-- can enter this in the Console to have output produced if not working here
```


# Use 'map' function -- map(your data inputs, function you want to apply to the inputs)

This one line of code replaces the for loops we did above 
```{r}
output = map(list_norm_samples, mean_and_sd)

output = map(list_norm_samples, summary)
```


# Create a Data Frame 

```{r}
listcol_df = 
  tibble(
    name = c("a", "b", "c", "d"),
    samp = list_norm_samples
  )
```


```{r}
listcol_df %>% pull(samp)

mean_and_sd(listcol_df$samp[[1]])
mean_and_sd(listcol_df$samp[[2]])
mean_and_sd(listcol_df$samp[[3]])

# Instead of those 3 simplify to one line of code with map 

map(listcol_df$samp, mean_and_sd)

listcol_df %>% 
  mutate(mean_sd = map(samp, mean_and_sd),
         median = map(samp, median)) %>% 
        unnest(mean_sd)

# In this code above - we start with a df with name and sample and we are adding a column here called mean_sd and a median column 
# The 'unnest()' function allows you to view the things 

```


# NSDUH Online data 

Read in the data
```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)
```

```{r}
nsduh_import = function(html, table_number, outcome_name){
    nsduh_html |> 
  html_table() |> 
  nth(table_number) |>
  slice(-1) |> 
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |>
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent),
    outcome = outcome_name  ) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
}
```


Import data using a 'for loop'

```{r}
table_input = list(1, 4, 5)
name_input = list("marj", "cocaine", "heroin")

output = vector("list", length = 3)

for (i in c(1:3)) {
 output[[i]] = nsduh_import(nsduh_html, table_input[[i]], name_input[[i]])
}

nsduh_df = bind_rows(output)
```


Lets do the same thing but make it better using maps!

```{r}
nsduh_import = function(html, table_number){
    nsduh_html |> 
  html_table() |> 
  nth(table_number) |>
  slice(-1) |> 
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |>
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
}

nsduh_df = 
  tibble(
    name = c("marj", "cocaine", "heroin"),
    number = c(1, 4, 5)
  ) %>% 
  mutate(
    table = map(number, nsduh_import, html = nsduh_html)) %>% 
  unnest(table)

## The goal of the code above is to keep track of the table name and tabe # in the data frame and then adding the whole dataset that corresponds to that as a listcol 

map(nsduh_df$number, nsduh_import, html = nsduh_html)
  
```

# Revist the Weather Dataset 

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

```{r}
weather_nest_df =
  weather_df %>% 
  nest(df = date:tmin)
```

Can I regress on tmax on tmin for each of these? 

```{r}
central_park_df = 
  weather_nest_df %>% 
  pull(df) %>% 
  nth(1)
```


Fit a linear regression for central park: 
```{r}
weather_lm = function(df) {
  lm(tmax ~ tmin, data = df)
  }


weather_lm(central_park_df)
```

# How can you fit separate linear models in each case? You have a few options...

Option 1: write a for loop 
```{r}
input_list = weather_nest_df %>% pull(df)
output = vector("list", length = 3)

for (i in 1:3) {
    output[[i]] = weather_lm(input_list[[i]]) 
}

weather_nest_df %>% 
  mutate(models = map(df, weather_lm))

```

