---
title: "Writing Functions"
output: github_document
---

```{r setup, include=FALSE}

library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuious.color = "viridis",
  ggplot2.continuious.fill = "viridis"
  
)

scale_color_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


```{r}
library(tidyverse)
library(rvest)
library(p8105.datasets)

```

Set seed for reproducibility: 
```{r}
set.seed(12345)
```

# Z score functions

Z scores subtract the mean and divide by the standard deviation

```{r}
x_vec = rnorm(20, mean = 5, sd = .3)
```

Compute Z score for x_vec: 

```{r}
(x_vec - mean(x_vec)) / sd(x_vec)
```

Lets create a Function to do this Z score instead!

```{r}
z_scores = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric")
  } else if (length(x) == 1) {
    stop("Z scores cannot be computed for length 1 vectors")
  }
  
  z = mean(x) / sd(x)
  
  z
}
```
if you tried to access this z variable it will not be in your global library - it will only exist inside this function 
-- good practice to specify these arguments to make LOUD error messages when your function does not work properly 


Check this function works: 
```{r}
z_scores(x_vec)
```


# Multiple outputs

Write a function that returns mean and st dev from a sample of numbers

```{r}
mean_and_sd = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric")
  } else if (length(x) == 1) {
    stop("Cannot be computed for length 1 vectors")
  }
  
  mean_x = mean(x)
  sd_x = sd(x)

   tibble(
    mean = mean_x, 
    sd = sd_x
  )
}
```

Check if this function works:

```{r}
mean_and_sd(x_vec)
```


# Start getting means an SDs

```{r}
x_vec = rnorm(n = 30, mean = 5, sd = .5)

tibble(
    mean = mean(x_vec), 
    sd = sd(x_vec)
  )
```

Lets write a funciton that uses 'n', a true mean, and true SD as inputs

```{r}
sim_mean_sd = function(n_obs, mu, sigma) {
  x_vec = rnorm(n = n_obs, mean = mu, sd = sigma)
  
tibble(
    mean = mean(x_vec), 
    sd = sd(x_vec)
  )
}
```
if i wanted to make this function have a contstant value for mean or sd i would specify this in the first line EX: sim_mean_sd = function(n_obs, mu = 2, sigma = .5)  
-- then you would only need to specify the n_obs for the function to work bc mu and sigma are preset 


Testing this sim_mean_sd function: 
```{r}
sim_mean_sd(n_obs = 30, mu = 5, sigma = .5)
sim_mean_sd(30, 5, .5)

sim_mean_sd(n_obs = 50, mu = 10, sigma = .5)
```
if you do not explicity name mu = 5 etc in the () then R assumes you are putting the numbers in the correct order you specified in the actual function 


# Revisit LOTR Data

```{r}
fellowship_ring = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "B3:D6") %>% 
  mutate(movie = "fellowship_ring")

lotr_load_and_tidy = function(path ="./data/LotR_Words.xlsx", cell_range, movie_name) {
  
movie_df = 
  readxl::read_excel(path, range = cell_range) %>% 
  mutate(movie = movie_name) %>% 
  janitor::clean_names() %>% 
  pivot_longer(
   female:male,
   names_to = "sex",
   values_to =  "words"
  ) %>% 
  select(movie, everything())
}

lotr_df = 
  bind_rows(
    lotr_load_and_tidy(cell_range = "B3:D6", movie_name = "fellowship_ring"),
    lotr_load_and_tidy(cell_range = "F3:H6", movie_name = "two_towers"),
    lotr_load_and_tidy(cell_range = "J3:L6", movie_name = "return_king"))

lotr_df
```


# Revisitng online NSDUH data 
```{r}
library(rvest)

nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)

data_marj = 
  nsduh_html |> 
  html_table() |> 
  nth(1) |>
  slice(-1) |> 
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |>
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))

view(data_marj)
```

Try to write a quick function to read in these online datatables from that html

```{r}
nsduh_import = function(html, table_number, outcome_name){
    nsduh_html |> 
  html_table() |> 
  nth(table_number) |>
  slice(-1) |> 
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |>
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent),
    outcome = outcome_name  ) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
}

nsduh_import(html = nsduh_html, table_number = 1, outcome_name = "marj")
nsduh_import(html = nsduh_html, table_number = 4, outcome_name = "cocaine")


```

