---
title: "Strings and Factors"
output: github_document
---

```{r setup, include=FALSE}

library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuious.color = "viridis",
  ggplot2.continuious.fill = "viridis"
  
)

scale_color_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


```{r}
library(rvest)
library(p8105.datasets)

```


## Strings and manipulations

```{r}

string_vec = c("my", "name", "is", "vicky")

str_detect(string_vec, "vicky")
str_detect(string_vec, "y")
str_detect(string_vec, "Vicky")

str_replace(string_vec, "vicky", "Victoria Mello")
```

```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "i think")
str_detect(string_vec, "^i think")
str_detect(string_vec, "i think$")

```
the ^ = at the beginning and the $ = at the end 


```{r}
string_vec = c(
  "Time for a Pumpkin Spice Latte!",
  "went to the #pumpkinpatch last weekend",
  "Pumpkin Pie is obviously the best pie",
  "SMASHING PUMPKINS -- LIVE IN CONCERT!!"
  )

str_detect(string_vec,"Pumpkin")

str_detect(string_vec,"[Pp]umpkin")
```
the [Pp] is how you make it look for both upper and lowercase 

```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "[0-9]")
str_detect(string_vec, "^[0-9]")
str_detect(string_vec, "^[0-9][a-z]")

```

```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11")
```
the . allows you to look for 711 separated by ANY SYMBOL (but not if theres no symbol)


```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, "\\[")
```
indicate that the thing you are looking for is a 'special character' with a double backslash \\


## Factors 

```{r}
vec_sex = c("male", "male", "female", "female")
vec_sex

as.numeric(vec_sex)

vec_sex = fct_relevel(vec_sex, "male")
vec_sex

as.numeric(vec_sex)

```


## NSDUH 

```{r}
nsduh_url = "https://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm" 

table_marj = 
  read_html(nsduh_url) %>% 
  html_table() %>% 
  first() %>% 
  slice(-1)
```

--> This is not a tidy dataset! We need to tidy this data imported from web before working w it...

```{r}
marj_df = 
  table_marj %>% 
  select(-contains("P Value")) %>% 
  pivot_longer(
      -State,
      names_to = "age_year",
      values_to = "percent"
  ) %>% 
  separate(age_year, into = c("age", "year"), "\\(") %>% 
   mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)) %>% 
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))

marj_df |>
  filter(age == "12-17") |> 
  mutate(State = fct_reorder(State, percent)) |> 
  ggplot(aes(x = State, y = percent, color = year)) + 
    geom_point() + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1))

```


## NYC Resturant Inspections dataset 

```{r}
data("rest_inspec")
rest_inspec
```

```{r}
rest_inspec |> 
  group_by(boro, grade) |> 
  summarize(n = n()) |> 
  pivot_wider(names_from = grade, values_from = n)
```

```{r}
rest_inspec =
  rest_inspec |>
  filter(grade %in% c("A", "B", "C"), boro != "Missing") |> 
  mutate(boro = str_to_title(boro))
```


Lets look at Pizza Places:

```{r}
rest_inspec %>% 
  filter(str_detect(dba, "PIZZA")) %>% 
  count(boro)

```

```{r}
rest_inspec %>% 
  filter(str_detect(dba, "PIZZA")) %>% 
  mutate(boro = fct_infreq(boro)) %>% 
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar()

```
--> Jeff likes fct_reoder bc it says that you want the average for the numeric variable to be the order they are in 
** IF you wanted to do something like rename Manhattan "The City" you would use fct_recode(boro, "The City" = Manhattan))
